---
title: "PS 1 SOLUTION SET"
author: Spring 2024
output: 
  pdf_document:
    includes:
      in_header: header.tex
---


There are usually many ways to arrive at complete and correct answers. The solution set shows you one way. As always, study the keys (the `.Rmd` code and `.pdf` output), and reach out with questions. 

```{r setup, include=TRUE}
## Set code chunk options
  knitr::opts_chunk$set(
    echo = TRUE,
    eval = TRUE,
    fig.align = 'center',
    message = FALSE,
    warning = FALSE
  )

knitr::knit_hooks$set(
  inline = function(x) {
       if(!is.numeric(x)){x}else{
         prettyNum(round(x, 2), big.mark=",")
       } 
  }
)
```

```{r}
## Load packages
  library(tidyverse)
  library(knitr)

## Read greenspace data  
  df = read_csv("https://dataverse.harvard.edu/api/access/datafile/6903364")

## clean
  gspace =
    df |>
    select(-1) |>
    mutate(
      ind_2021f =
        fct_relevel( # for the freq table
          indicator_2021,
          "Very High",
          "High",
          "Moderate",
          "Low",
          "Very Low",
          "Exceptionally Low",
        )
    ) 
```

1. The code and output files (`Hart PSet2.Rmd` and `Hart PSet2.pdf`) are available in my PSet2 repository on `GitHub`: <https://github.com/austin-hart-pols/PSet2>.

2. See the corrected code below. I identified 7 errors in total.

```{r, eval = FALSE}
  library(tidyverse) 
  library(knitr)  # 1. load knitr for kable

# open my data
  gspace = read_csv('greenspace_data_share.csv') # 2. quote filename
  
# summarize average urban greenspace by region
  table =
    gspace |> 
    group_by(Major_Geo_Region) |> # 3. add pipe
    summarise(
      obs = n(), # 4. comma at end
      avg = mean(annual_avg_2020), #5. fix var name
      `weighted avg` = mean(annual_weight_avg_2020) # 6. fix output name
  )
# output as table
  kable(table, digits = 1) # 7. correct object is table
```

3. The Global Greenspace Indicator Data includes records for `r nrow(gspace)` urban areas.

4. The table below shows the Greenspace Indicator score for urban areas in 2021. Note that only 6\% of urban areas are scored high to exceptionally high on this scale while almost three-quarters (roughly 75\%) are scored low to exceptionally low. "Low" is the most common score (38\%), and about one in five urban areas have "Moderate" greenspace.

```{r}
  gspace |>
    count(ind_2021f, name = "Freq") |>
    mutate(
      Per = Freq / sum(Freq) * 100
    ) |>
    rename(`Greenspace (2021)` = ind_2021f) |>
    kable(digits = 1L, caption = 'Greenspace in Urban Areas, 2021')
```

5. See below.  
    a. `r  gspace |> filter(indicator_2015 %in% c('High', 'Very High')) |> nrow()` urban areas scored High or above in 2015.  
    b. `r gspace |> filter(if_any(starts_with('indicator'),  ~ .x == "Exceptionally Low")) |> nrow()` areas recorded a score of exceptionally low at some point from 2010 to 2021.  
    c. Among `r gspace |> filter(Climate_region == 'Arid') |> nrow()` urban areas in arid climates, `r gspace |> filter(Climate_region == 'Arid', annual_weight_avg_2020 > annual_weight_avg_2010) |> nrow()` recorded increases in NDVI greenspace (annual, weighted average) from 2010 to 2020.  
    
6. The table below shows regional differences in the number of urban areas that lost greenspace (NDVI, annual average) from 2010 to 2021. Of the 128 urban areas where the NDVI score declined, a strong plurality (37\%) come from Europe.

```{r}
  gspace |>
    mutate(
      less_green = annual_avg_2010 > annual_avg_2021
    ) |>
    count(less_green, Major_Geo_Region) |>
    pivot_wider(
      names_from = less_green,
      values_from = n,
      values_fill = 0
    ) |>
    mutate(Percent = `TRUE` / (`FALSE` + `TRUE`) * 100) |>
    select(1, 3, 5) |>
    rename(
      Continent = Major_Geo_Region,
      Freq = `TRUE`
    ) |>
    kable(digits = 1L, caption = 'Loss of Greenspace by Region, 2010-21')
```  

7. See the histogram below.

```{r, fig.dim=c(3,2.5), fig.cap = 'Growth and loss of urban greenspace, 2010 to 2021'}
  gspace |>
    mutate(
      change = annual_avg_2021 - annual_avg_2010
    ) |>
    ggplot(aes(x = change, fill = change > 0)) +
    geom_histogram(
      color = 'white', 
      boundary = 0,
      closed = 'left'
    ) +
    labs(
      x = "Change in NDVI, 2010 to 2021",
      y = "Urban areas"
    ) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    scale_fill_manual(values = c('orange2','seagreen4')) +
    theme_classic() +
    theme(legend.position = 'none')
```

8. The scatter plot below shows the response with the additional options for the bonus point.

```{r, fig.dim=c(3,2.5), fig.cap = 'Changes in urban greenspace'}
  gspace |>
    ggplot(aes(x = annual_weight_avg_2010, y = annual_weight_avg_2021)) +
    geom_abline(intercept = 0, slope = 1) +
    geom_point(
      aes(color = annual_weight_avg_2021 > annual_weight_avg_2010),
      alpha = .5
    ) +
    scale_y_continuous(limits = c(0, 0.75)) +
    scale_x_continuous(limits = c(0, 0.75)) +
    scale_color_manual(values = c('orange2','seagreen4')) +
    labs(
      y = 'NDVI 2021 (weighted)',
      x = 'NDVI 2010 (weighted)'
    ) +
    theme_classic() +
    theme(legend.position = 'none')
```
